---
title: "Labs"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)

```

# Covid Cases



```{r}
# Lab 2 Data
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)
library(dplyr)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'

covid = read_csv(url)

pop = PopulationEstimates <- read_excel("~/Downloads/PopulationEstimates.xls", 
                                        skip = 2)
 #Set up data by county/create a New Cases column
data = covid %>%
  filter(state == "California") %>%
  group_by(county) %>%
  mutate(newCases = cases - lag(cases)) %>%
  ungroup()

today = data %>% filter(date == max(date))


# Calculate total California Cases for end analysis 
sum(today$cases)
sum(today$newCases)


# Make Tables - Cumulative/New Cases
most_new_cases = today %>% 
  slice_max(newCases, n = 5) %>%
  select(county, newCases)

most_cumulative_cases = data %>%
  filter(date == max(date)) %>%
  slice_max(cases, n = 5) %>%
  select(county, cases)


# Refine Tables - cumulative/new cases
library(kableExtra)
knitr::kable(most_new_cases,
             caption = "California Counties: Most New Cases", 
             col.names = c("County", "New Cases"), 
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)

knitr::kable(most_cumulative_cases,
             caption = "California Counties: Most Cumulative Cases",
             col.names = c("County", "Cumulative Cases"),
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)




# Join population to covid data/filter to include only last 14 days 
perCap = pop %>%
  select(fips = "FIPStxt", state = "State", pop2019 = "POP_ESTIMATE_2019") %>%
  right_join(data, by = "fips") %>%
  mutate(cumulative_cases_percapita = (cases/pop2019), new_cases_percapita = (newCases/pop2019)) %>%
  filter(date >= max(date) - 13)


# Make Percapita Tables
cum_per_cap = perCap %>%
  filter(date == max(date)) %>%
  slice_max(cases, n = 5) %>%
  select(county, cumulative_cases_percapita)

new_per_cap = perCap %>%
  filter(date == max(date)) %>%
  slice_max(cases, n = 5) %>%
  select(county, new_cases_percapita)

# Refine Tables - Per Capita 

knitr::kable(cum_per_cap,
             caption = "California Counties: Most Cumulative Cases Per Capita", 
             col.names = c("County", "Cumulative Cases Per Capita"), 
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)



knitr::kable(new_per_cap,
             caption = "California Counties: Most New Cases Per Capita",
             col.names = c("County", "New Cases Per Capita"),
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)


# Calculate number of new cases in last 14 days per 100000 people/number of safe counties 
safe = perCap %>%
  group_by(county, pop2019) %>%
  summarise(total_newCases = sum(newCases, na.rm = TRUE)) %>% 
  mutate(total_safecounties = total_newCases/(pop2019/100000)) %>% 
  filter(total_safecounties < 100)


# Facet plot 1
data2 = covid %>%
  filter(state %in% c("New York","California", "Louisiana", "Florida")) %>%
  group_by(state, date) %>%
  summarize(cases = sum(cases)) %>% 
  ungroup() %>% 
  group_by(state) %>% 
  mutate(newCases = cases - lag(cases),
         roll7 = zoo::rollmean(newCases, 7, fill = NA, align = 'right')) %>%
  ungroup() %>%
  filter(newCases > 0)

ggplot(data = data2, aes(x = date)) +
  geom_col(aes(y = newCases), col = NA, fill = "red") +
  geom_line(aes(y = roll7), col = "darkred", size = 1) +
  ggthemes::theme_wsj() +
  labs(title = paste("New Cases")) +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(size = 14, face = 'bold')) +
  facet_grid(~state, scales = "free_y")


data3 = pop %>%
  select(pop2019 = "POP_ESTIMATE_2019", areaname = "Area_Name") %>%
  right_join(data2, by = c("areaname" = "state")) %>%
  mutate(cases_per_capita = (newCases/pop2019), roll7 = zoo::rollmean(cases_per_capita, 7, fill = NA, align = 'right')) %>%
  ungroup() %>%
  filter(cases_per_capita > 0)

ggplot(data = data3, aes(x = date)) +
  geom_col(aes(y = cases_per_capita), col = NA, fill = "lightblue") +
  geom_line(aes(y = roll7), col = "blue", size = 1) +
  ggthemes::theme_wsj() +
  labs(title = paste("New Cases Per Capita")) +
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill = "white"),
        plot.title = element_text(size = 14, face = 'bold')) +
  facet_grid(~areaname, scales = "free_y")


```

# Cities and Borders


```{r}


library(tidyverse)
library(sf)
library(units)

# Data
library(USAboundaries)

library(rnaturalearth)

# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)

library(ggthemes) 

# Question 1.1 -Define a Projection 
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'


# Question 1.2 - Get USA state boundaries 
data = USAboundaries::us_states(resolution = "low") %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) %>%
  st_transform(eqdc)

# Question 1.3 
country = rnaturalearth::countries110

country_sf = st_as_sf(country) %>%
  filter(admin %in% c("United States of America", "Mexico", "Canada")) %>%
  st_transform(eqdc)

# Question 1.4- City locations

cities = readr::read_csv('~/Documents/github/exercise-04/data/uscities.csv')

cities_sf = data.frame(y = c(-121.7608, -119.7202), x = c(34.4285, 37.6861))
cities_sf = st_as_sf(cities, coords = c("lng", "lat"), crs = 4326) %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii"))

cities_eqdc = st_transform(cities_sf, eqdc) %>% 
  select(city, state_name, population)


# Union and Combine States
usa_geom2 = data$geometry

ca_u_ml = st_union(usa_geom2)  %>%
  st_cast("MULTILINESTRING")

ca_c_ml = st_combine(usa_geom2)  %>%
  st_cast("MULTILINESTRING")


# Distance to US Border
cities_eqdc = cities_eqdc %>%
  mutate(dist_border = st_distance(cities_eqdc, ca_u_ml)) %>%
  mutate(dist_border = units::set_units(dist_border, "km"))
# Distance to state borders
cities_eqdc = cities_eqdc %>%
  mutate(dist_state = st_distance(cities_eqdc, ca_c_ml)) %>%
  mutate(dist_state = units::set_units(dist_state, "km"))
# Distance to Canada
cities_eqdc = cities_eqdc %>%
  mutate(dist_Canada = st_distance(cities_eqdc, filter(country_sf, admin == "Canada"))) %>%
  mutate(dist_Canada = units::set_units(dist_Canada, "km"))
# Distance to Mexico
cities_eqdc = cities_eqdc %>%
  mutate(dist_Mexico = st_distance(cities_eqdc, filter(country_sf, admin == "Mexico"))) %>%
  mutate(dist_Mexico = units::set_units(dist_Mexico, "km"))


# Distances:
# Cities to US Border 
cities_eqdc %>% select(city,state_name, dist_border) %>% slice_max(dist_border, n = 5) %>% 
  st_drop_geometry() %>% 
  knitr::kable(caption = "Cities Farthest from US Border",
               col.names = c("City", "State", "Distance to US Border"),
               format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)


# Cities to states
cities_eqdc %>% select(city,state_name, dist_state) %>% slice_max(dist_state, n = 5) %>% 
  st_drop_geometry() %>% 
  knitr::kable(caption = "Cities Farthest from State Boundaries",
               col.names = c("City", "State", "Distance to State Boundary"),
               format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)


# Distance to Mexico
cities_eqdc %>% select(city,state_name, dist_Mexico) %>% slice_max(dist_Mexico, n = 5) %>% 
  st_drop_geometry() %>% 
  knitr::kable(caption = "Cities Farthest from Mexican Border",
               col.names = c("City", "State", "Distance to Mexican Border"),
               format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14) 



# Distance to Canada
cities_eqdc %>% select(city,state_name, dist_Canada) %>% slice_max(dist_Canada, n = 5) %>% 
  st_drop_geometry() %>% 
  knitr::kable(caption = "Cities Farthest from Canadian Border",
               col.names = c("City", "State", "Distance to Canadian Border"),
               format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14) 



# 3.1 - 10 Largest Cities

big_cities2 = cities_eqdc %>%
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = country_sf) +
  geom_sf(data = data, lty = "dashed", size = .5) +
  geom_sf(data = big_cities2, size = 1) +
  ggrepel::geom_label_repel(
    data = big_cities2,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "Most Populous US Cities",
       x = " ",
       y = " ") +
  theme_economist()

# 3.2- City Distance to Border 
farthest_cities = cities_eqdc %>%
  slice_max(dist_border, n = 5)

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = cities_eqdc, aes(col = as.numeric(dist_border)), size = 0.1) +
  geom_sf(data = farthest_cities, size = 0.5, color = "red") +
  scale_color_gradient(low = "gray", high = "blue", name = "Distance (km)") +
  ggrepel::geom_label_repel(
    data = farthest_cities,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "City Distance from National Border & Farthest Cities",
       x = " ",
       y = " ") + 
  theme_economist() 


# 3.3 - City Distance from Nearest State 
farthest_cities_state = cities_eqdc %>%
  slice_max(dist_state, n = 5)

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = farthest_cities_state, size = 1, color = "red") +
  geom_sf(data = cities_eqdc, aes(col = as.numeric(dist_state)), size = 0.1) +
  scale_color_gradient(low = "gray", high = "blue", name = "Distance (km)") +
  ggrepel::geom_label_repel(
    data = farthest_cities_state,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "City Distance from Nearest State Border & Farthest Cities",
       x = " ",
       y = " ") + 
  theme_economist() 


# 3.4- Equidistance Boundary from Mexico and Canada 

cities_eqdc = cities_eqdc %>%
  mutate(difference = abs(dist_Canada - dist_Mexico)) 


big_cities3 = cities_eqdc %>%
  filter(as.numeric(difference) < 100 ) %>%
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = cities_eqdc, aes(col = as.numeric(difference)), size = 0.1) + 
  gghighlight::gghighlight(as.numeric(difference) < 100) +
  geom_sf(data = big_cities3, size = 0.5, color = "red") +
  scale_color_gradient(low = "gray", high = "blue", name = "Distance (km)") +
  ggrepel::geom_label_repel(
    data = big_cities3,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "Cities Equidistant from Canadian and Mexican Borders (Within 100 km)",
       subtitle = "And Most Populous Cities",
       x = " ",
       y = " ") + 
  theme_economist() 



# 4.1 - Quantifying Border Zone 
total_pop = cities_eqdc %>%
  select(population) %>%
  summarise(sum(population))

danger_zone = cities_eqdc %>%
  mutate(total_pop = sum(population)) %>%
  filter(as.numeric(dist_border) <= 160) %>%
  summarise(number_cities = n(), sum_pop = sum(population), percent = 100 * (sum_pop/total_pop[1])) %>%
  st_drop_geometry() 

knitr::kable(danger_zone, caption = "Cities Within 100 Miles of State Border",
             col.names = c("Number of Cities", "Number of People", "Percent of Population"),
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14) 


# 4.2 - Mapping Border Zone 
highest_pop_dangerzone = cities_eqdc %>%
  filter(as.numeric(dist_border) <= 160) %>%
  select(city, population) %>%
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = cities_eqdc, aes(col = as.numeric(dist_border)), size = 0.1) +
  gghighlight::gghighlight(as.numeric(cities_eqdc$dist_border) <= 160) +
  scale_color_gradient(low = "orange", high = "dark red", name = "Distance (km)") +
  geom_sf(data = highest_pop_dangerzone, size = 1) +
  ggrepel::geom_label_repel(
    data = highest_pop_dangerzone,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "Cities in 100 Mile Zone & Most Populous Cities",
       x = " ",
       y = " ") + 
  theme_economist() 
######################
##Extra Credit
highest_pop_state = cities_eqdc %>%
  filter(as.numeric(dist_border) <= 160) %>%
  select(city, state_name, population) %>%
  group_by(state_name) %>%
  filter(population == max(population))

ggplot() +
  geom_sf(data = data) +
  geom_sf(data = cities_eqdc, aes(col = as.numeric(dist_border)), size = 0.1) +
  gghighlight::gghighlight(as.numeric(cities_eqdc$dist_border) <= 160) +
  scale_color_gradient(low = "orange", high = "dark red", name = "Distance (km)") +
  geom_sf(data = highest_pop_state, size = 1) +
  ggrepel::geom_label_repel(
    data = highest_pop_state,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 2) +
  labs(title = "Cities in 100 Mile Zone & Most Populous City Per State",
       x = " ",
       y = " ") + 
  theme_economist() 


```

# Tessallations

```{r}
library(tidyverse)
library(sf)
library(USAboundaries)
library(rmapshaper)
library(readxl)
library(gghighlight)
library(leaflet)
library(leafpop)


# Step 1.1 - Get CONUS & Simplify 
conus = USAboundaries::us_counties() %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) %>%
  st_transform(5070)

conus_simp = ms_simplify(conus, keep = 0.05)

conuspts = mapview::npts(conus)
simppts = mapview::npts(conus_simp)
diff = conuspts - simppts



# Step 1.2 - Centroids 

county_centroid = st_centroid(conus_simp) %>%
  st_combine() %>%
  st_cast("MULTIPOINT")

# Step 1.3 - 1.5: Make Tessalations 

# Voroni Tessellation 
v_grid = st_voronoi(county_centroid) %>%
  st_cast() %>%
  st_as_sf %>%
  mutate(id = 1:n())


# Triangulated Tessalation
t_grid = st_triangulate(county_centroid) %>%
  st_cast() %>%
  st_as_sf() %>%
  mutate(id = 1:n())


# Gridded Coverage: n = 70
sq_grid = st_make_grid(conus_simp, n = 70) %>%
  st_as_sf() %>%
  mutate(id = 1:n())


# Hexagonal Coverage: n = 70
hex_grid = st_make_grid(conus_simp, n = 70, square = FALSE) %>%
  st_as_sf() %>%
  mutate(id = 1:n())


# 1.6 - Plot

plot_tess = function(data, title)
{ggplot() + 
    geom_sf(data = data, fill = "white", col = "navy", size = .2) +   
    theme_void() +
    labs(title = title, caption = paste("This tesselation has:", nrow(data), "tiles" )) +
    theme(plot.title = element_text(hjust = .5, color =  "black", face = "bold"))}

# Original
plot_tess(data = conus_simp, "Original County Data")


# Voroni
v_grid = st_intersection(v_grid, st_union(conus_simp))
plot_tess(v_grid, "Voronoi Coverage") +
  geom_sf(data = county_centroid, col = "lightgreen", size = 0.1)

# Triangulated

t_grid = st_intersection(t_grid, st_union(conus_simp))
plot_tess(t_grid, "Triangulated Coverage") +
  geom_sf(data = county_centroid, col = "lightgreen", size = 0.1)

# Gridded

plot_tess(sq_grid, "Square Coverage")


# Hexagonal

plot_tess(hex_grid, "Hexagonal Coverage")




# Question 2 - Still need to comment on features
tess_summary = function(sf_object, character_string){
  area = st_area(sf_object, na.rm = FALSE, weights = FALSE) %>%
    units::set_units("km2") %>%
    units::drop_units() 
  data.frame(type = character_string,
             features = nrow(sf_object),
             mean_area = mean(area),
             standard_deviation = sd(area),
             total_area = sum(area))
} 

table_summary = bind_rows(
  tess_summary(conus_simp, "Original"),
  tess_summary(v_grid, "Voroni"),
  tess_summary(t_grid, "Triangulation"),
  tess_summary(sq_grid, "Gridded"),
  tess_summary(hex_grid, "Hexagons"))

knitr::kable(table_summary,
             caption = "Tesselation Characteristics", 
             col.names = c("Type", "Number of Features", "Mean Area", "Standard Deviation (km^2)","Total Area (km^2)"), 
             format.args = list(big.mark = ",")) %>%
  kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)


# Question 3

library(readxl)
NID2019 =read_excel("~/Downloads/NID2019_U.xlsx")


dams_sf = NID2019 %>%
  filter(!is.na(LONGITUDE), !is.na(LATITUDE)) %>%
  st_as_sf(coords = c('LONGITUDE', 'LATITUDE'), crs = 4326) %>%
  st_transform(5070)


point_in_polygon = function(points, polygon, id){
  st_join(polygon, points) %>%
    st_drop_geometry() %>%
    count(.data[[id]]) %>%
    setNames(c(id, "n")) %>%
    left_join(polygon, by = id) %>%
    st_as_sf()
}

orig_pip = point_in_polygon(dams_sf, conus_simp, "geoid")

# Voroni
voroni_pip = point_in_polygon(dams_sf, v_grid, "id")

# Triangulation
tri_pip = point_in_polygon(dams_sf, t_grid, "id")

# Grid
grid_pip = point_in_polygon(dams_sf, sq_grid, "id")

# Hexagon Grid
hex_pip = point_in_polygon(dams_sf, hex_grid, "id")

# Plot Function 

plot_tess_pip = function(sf_tess_object, title)
{ggplot() + 
    geom_sf(data = sf_tess_object, aes(fill = n), col = NA) +
    scale_fill_viridis_c(name = " ") + 
    theme_void() + 
    theme(plot.title = element_text(face = "bold", color = "black", hjust = .5, size = 24)) +
    labs(title = paste("Dam Locations:", title),
         caption = paste0(sum(sf_tess_object$n), " Dam Locations Represented"))}
# Plot
plot_tess_pip(orig_pip, "Original")
plot_tess_pip(voroni_pip, "Voroni")
plot_tess_pip(tri_pip, "Triangulation")
plot_tess_pip(grid_pip, "Square Grid")
plot_tess_pip(hex_pip, "Hexagon Grid")


# Flood Control Dams
flood_control =  dams_sf %>%
  filter(grepl("C", PURPOSES)) 

flood_pip = point_in_polygon(flood_control, v_grid, "id")

# Fire Protection Dams
fire_control =  dams_sf %>%
  filter(grepl("P", PURPOSES)) 

fire_pip = point_in_polygon(fire_control, v_grid, "id")

# Water Supply

water_supply =  dams_sf %>%
  filter(grepl("S", PURPOSES))

water_pip = point_in_polygon(water_supply, v_grid, "id")

# Fish and Wildlife

fish_wildlife =  dams_sf %>%
  filter(grepl("F", PURPOSES))

fish_wildlife_pip = point_in_polygon(fish_wildlife, v_grid, "id")

# Plots

# Flood Control 
plot_tess_pip(flood_pip, "Flood Control") +
  gghighlight::gghighlight(n > mean(n) + 1) 
plot_tess_pip(fire_pip, "Fire Control") +
  gghighlight::gghighlight(n > mean(n) + 1) 
plot_tess_pip(water_pip, "Water Supply") +
  gghighlight::gghighlight(n > mean(n) + 1) 
# Fish and Wildlife
plot_tess_pip(fish_wildlife_pip, "Fish and Wildlife") +
  gghighlight::gghighlight(n > mean(n) + 1) 
```
# Palo Flood Landsat Bands

```{r}
library(readr)
library(dplyr)
library(raster) 
library(tidyverse) 
library(getlandsat) 
library(sf)
library(mapview) 
library(whitebox)
library(USAboundaries)
library(kableExtra)

################
PaloBoundary = read_csv("~/Documents/github/176A-Lab01/data/uscities.csv") %>%
  filter(city == "Palo") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(5070) %>%
  st_buffer(10000) %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_as_sf()

bb = read_csv("~/Documents/github/176A-Lab01/data/uscities.csv") %>%
  filter(city == "Palo") %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(5070) %>%
  st_buffer(5000) %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_as_sf()


bbwgs = PaloBoundary %>%
  st_transform(4326)
PaloBoundaryWgs = st_bbox(bbwgs)
meta = read_csv("~/Documents/github/176A-Lab01/data/palofloodscene.csv")




  
  
  
  
  
  
  files = lsat_scene_files("https://s3-us-west-2.amazonaws.com/landsat-pds/L8/025/031/LC80250312016270LGN00/index.html") %>% 
    filter(grepl(paste0("B", 1:6, ".TIF$", collapse = "|"), file))  %>% 
    arrange(file) %>% 
    pull(file)
  
  st = sapply(files, lsat_image)
  s = stack(st) %>% 
    setNames(c(paste0("band", 1:6)))  

  
  cropper = bbwgs %>% 
    st_transform(crs(s))
  r = crop(s, cropper)
  dim(r)  
  
  par(mfrow = c(2,1))
  plotRGB(r, r = 4, g = 3, b = 2, stretch = "lin")
  plotRGB(r, r = 4, g = 3, b = 2, stretch = "hist")  
  
  plotRGB(r, r = 5, g = 5, b = 3, stretch = "lin")
  plotRGB(r, r = 5, g = 5, b = 3, stretch = "hist")
  
  plotRGB(r, r = 5, g = 6, b = 4, stretch = "lin")
  plotRGB(r, r = 5, g = 6, b = 4, stretch = "hist")
  
  plotRGB(r, r = 5, g = 3, b = 2, stretch = "lin")
  plotRGB(r, r = 5, g = 3, b = 2, stretch = "hist")

  #4.1
  
  palette = colorRampPalette(c("blue", "white", "red"))
  
  
  ndvi = (r$band5 - r$band4) / (r$band5 + r$band4)
  plot(ndvi, col = palette(256))  
  
  ndwi = (r$band3 - r$band5) / (r$band3 + r$band5)
  plot(ndwi, col = palette(256))
  
  mndwi = (r$band3 - r$band6) / (r$band3 + r$band6)
  plot(mndwi, col = palette(256))

  
  wri = (r$band3 + r$band4) / (r$band5 + r$band6)
  plot(wri, col = palette(256))  
  
  
  swi = (1 / sqrt(r$band2 - r$band6))
  plot(swi, col = palette(256))
  
  stack_indices = raster::stack(ndvi, ndwi, mndwi, wri, swi) %>% 
    setNames(c("NDVI", "NDWI", "MNDWI", "WRI", "SWI"))
  plot(stack_indices)
  
  
  #4.2 
  thresholding1 = function(x){ifelse(x <= 0,1,NA)}
  thresholding2 = function(x){ifelse(x >= 0,1,NA)}
  thresholding3 = function(x){ifelse(x >= 0,1,NA)}
  thresholding4 = function(x){ifelse(x >= 1, 1,NA)}
  thresholding5 = function(x){ifelse(x <= 5,1,NA)}
  
  flood1 = calc(ndvi, thresholding1)
  plot(flood1, col = "blue")

  
  flood2 = calc(ndwi, thresholding2)
  plot(flood2, col = "blue")  
  
  flood3 = calc(mndwi, thresholding3)
  plot(flood3, col = "blue")
  
  flood4 = calc(wri, thresholding4)
  plot(flood4, col = "blue")

  flood5 = calc(swi, thresholding5)
  plot(flood5, col = "blue")  
  
  floods = raster::stack(flood1, flood2, flood3, flood4, flood5) %>% 
    setNames(c("ndvi", "ndwi", "mndwi", "wri", "swi"))
  

  plot(floods, col = blues9)

  
  
  
  
  #5.1
 
  set.seed(09062020)
  
  #5.2 
  r_values = getValues(r) %>% 
    na.omit()
  
 
  k_12 = kmeans(r_values, 12)
  k_6 = kmeans(r_values, 6)
  k_4 = kmeans(r_values, 4)
  
  
  kmeans_raster = stack_indices$NDVI
  kmeans_raster6 = stack_indices$NDVI
  kmeans_raster4 = stack_indices$NDVI
  
  
  values(kmeans_raster) = k_12$cluster
  values(kmeans_raster6) = k_6$cluster
  values(kmeans_raster4) = k_4$cluster
  

  plot(kmeans_raster)

  plot(kmeans_raster6)  

  plot(kmeans_raster4)
 
  
   #5.3 
  
  floods_values = values(flood1)
  k_12_values = values(kmeans_raster)
  kmeans_table = table(floods_values, k_12_values)
  
  idx = which.max(kmeans_table)
  
  kmeans_raster[kmeans_raster != idx] = 0
  
  kmeans_stack = stack(flood1, flood2, flood3, flood4, flood5, kmeans_raster)
  
  plot(kmeans_stack)
  ##6
  plot_sum = calc(kmeans_stack, fun = sum)
  
  floods_stats = cellStats(kmeans_stack, stat = sum)*res(kmeans_stack)^2/1e6
  
  kable(floods_stats, col.names = "Area (m^2)")
  
  
  plot(plot_sum, col = blues9)
  
  
  plot_sum[plot_sum == 0] = NA
  mapview(plot_sum)
  
  new_aoi1 = st_point(c(-91.78946, 42.06303)) %>% 
    st_sfc(crs = 4326) %>% 
    st_as_sf() %>% 
    st_transform(st_crs(kmeans_stack))
  
  new_aoi2 = st_point(c(-91.78482, 42.06245)) %>% 
    st_sfc(crs = 4326) %>% 
    st_as_sf() %>% 
    st_transform(st_crs(kmeans_stack))
  
  f_values = raster::extract(kmeans_stack, new_aoi1) 
  f_values_plot = new_aoi1 %>% 
    mutate(f_values = f_values)  
```
# Santa Barbara Flood

```{r cars}


# Data libraries
library(osmdata)   
library(elevatr) 

library(tidyverse)
library(sf)       
library(raster)   
library(fasterize) 
library(whitebox)  
library(gifski)
library(knitr)

# Data libraries
library(osmdata)   
library(elevatr) 
basin = read_sf("https://labs.waterdata.usgs.gov/api/nldi/linked-data/nwissite/USGS-11119750/basin")

elev = elevatr::get_elev_raster(basin, z = 13) %>% 
  crop(basin) %>% 
  mask(basin)

elev = elev * 3.281


writeRaster(elev, "~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif", overwrite = T)

elev_raster = raster("~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif")


bb = st_bbox(basin) %>% 
  st_as_sfc() %>% 
  st_transform(4326)

build = osmdata::opq(bb) %>% 
  add_osm_feature(key = "building") %>% 
  osmdata_sf()

building_point = build$osm_polygons %>% 
  st_intersection(basin) %>% 
  st_transform(crs(basin)) %>% 
  st_centroid()

building_poly = build$osm_polygons %>% 
  st_intersection(basin) %>% 
  st_transform(crs(basin))

rail = opq(bb) %>% 
  add_osm_feature(key = "railway", value = "station") %>% 
  osmdata_sf()

railway = rail$osm_points %>% 
  st_intersection(basin) %>% 
  st_transform(crs(basin))

osm = opq(bb) %>% 
  add_osm_feature(key = "waterway", value = "stream") %>% 
  osmdata_sf()

stream = osm$osm_lines %>% 
  st_intersection(basin) %>% 
  st_transform(crs(basin))

hillshade = raster("~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif")

plot(hillshade, col = gray.colors(256, alpha = .5), main = "Hillshade Basin and Stream", legend = F, box = F)
plot(stream, add = T, col = "blue")
plot(basin, add = T)
```

# Code that would not work because wbt_whiteshade couldn't create a hillshade file
wbt_hillshade ( "~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif","~/Documents/github/tiagoschwartz.io/Mission-Creek-Hillshade.tif")


hillshade = raster("~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif")

plot(hillshade, col = gray.colors(256, alpha = .5), main = "Hillshade Basin and Stream", legend = F, box = F)
plot(stream, add = T, col = "blue")
plot(basin, add = T)

stream_buffer = stream %>% 
  st_transform(5070) %>% 
  st_buffer(10) %>% 
  st_transform(4326)

stream_raster = fasterize::fasterize(stream_buffer, elev_raster)

writeRaster(stream_raster, "~/Documents/github/tiagoschwartz.io/stream-raster.tif", overwrite = T)

wbt_breach_depressions("~/Documents/github/tiagoschwartz.io/Mission-Creek-Elev.tif", "~/Documents/github/tiagoschwartz.io/breach-depressions.tif")

wbt_elevation_above_stream("~/Documents/github/tiagoschwartz.io/breach-depressions.tif", "~/Documents/github/tiagoschwartz.io/stream-raster.tif", "~/Documents/github/tiagoschwartz.io/hand-raster.tif")

hand_rast = raster("~/Documents/github/tiagoschwartz.io/hand-raster.tif")
flood_raster = raster("~/Documents/github/tiagoschwartz.io/stream-raster.tif")

offset_raster = hand_rast + 3.69

offset_raster[flood_raster == 1] = 0

writeRaster(offset_raster, ".~/Documents/github/tiagoschwartz.io/offset_basin.tif", overwrite = T)


offset_raster = raster("~/Documents/github/tiagoschwartz.io/offset_basin.tif")

offsetraster1 = offset_raster

offset_raster[offset_raster > 10.02] = NA

plot(hillshade, col = gray.colors(256, alpha = .5), main = "Hillshade Basin and Flood", legend = F, box =F)
plot(offset_raster, add = T, col = rev(blues9), legend = F)
plot(basin, add = T)
plot(railway, col = "green", cex = 1, pch = 16, add = T)

building_points = ifelse(!is.na(raster::extract(offset_raster, building_point)), "red", "black")

plot(hillshade, col = gray.colors(256, alpha = .5), main = paste("2017 Mission Creek Basin Flood Affected", sum(building_points == "red"), "Buildings"), legend = F, box =F)
plot(offset_raster, add = T, col = rev(blues9), legend = F, box = F)
plot(basin, add = T)
plot(railway, col = "green", cex = 1, pch = 16, add = T, box = F)
plot(building_point, add = T, col = building_points, cex = .08, pch = 16)
